Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-04-10 Chapter-1/1-Theory.tex
s

Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories as a function of increasing neutrino energy: quasi elastic (QE), resonant and deep inelastic scattering (DIS). All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into resonant scattering and the energy  where resonance scattering  transitions in to DIS. 
 Neutrino and antineutrino QE charge current scattering refers to the process $\nu_\mu n \rightarrow \mu^- p$ and $\bar\nu_\mu p \rightarrow \mu^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction and the target nucleus remains at ground state. %of a neutrino (or antineutrino) with a nucleon in the target material.  
 Resonant scattering refers to an inelastic collision producing a nucleon excited state ($\Delta$,N*); the excited resonance quickly decays, most often to a nucleon and single-pion final state. DIS refers to the head-on collision between the neutrino and a parton inside the nucleon, producing hadronization and subsequent  abundant production of mesons and nucleons.  In addition to such interactions between the neutrino and a single component of the nucleus, neutrinos can also interact with the nucleus as a whole, albeit more rarely, a well documented process called coherent meson production scattering \cite{Vilain1993}; the signature of such process is the production of a distinctly forward-scattered single meson final state, most often a pion.  This simple picture of neutrino interactions works rather well for scattering 
off of light targets, such as the H$_2$ and D$_2$ of bubble chamber experiments \cite{RevModPhys.84.1307}, but the complexity of the nuclear structure for heavier nuclei such as argon complicates this model.

@@

Nuclear effects can potentially affect the neutrino event rates, nucleon emission, neutrino energy reconstruction, and neutrino/antineutrino ratios, carrying deep implications for oscillation experiments. 
Even in the case of ``simple" QE scattering  nuclear effects can impact the size and shape of the cross section as well as the final state kinematics and topology. Intra-nuclear hadron rescattering and correlation effects between the target nucleons can cause the ejection of additional nucleons in the final state. In case of resonant and DIS scattering, the hadronic interactions of meson and nucleons produced in the decay of the resonance or during hadronization  complicate this picture even more.
A large source of uncertainty in modeling  nuclear effects in neutrino interactions come from mesons interactions (and re-interactions) in the nucleus, e.g., pion re-scattering, charge exchange, and absorption.

A renewed interest for neutrino cross section measurements surged in recent year, along with lively discussion on the data reporting; the historical method of reporting the neutrino cross section as a function of the neutrino energy or momentum transferred shakes under the weight of its dependency to the chosen nuclear model.   On one hand, correcting for nuclear effects in neutrino interaction can introduce unwanted sources of uncertainty and model dependence due to the mis-modeling of the meson interactions. On the other, avoiding this correction however makes a comparison between neutrino interactions on different target nuclei extremely difficult.

Data on neutrino scattering off many different nuclei are available for both charged current (CC) and neutral current (NC) channels, as summarized here \cite{RevModPhys.84.1307}. A summary of is reported in Figure \textcolor{red}{PLOT} where the (NUANCE) [46] event generator provides a theoretical comparator.

>>>>

Historically, neutrino interactions with the nucleus in the GeV region are divided into three categories as a function of increasing neutrino energy: quasi elastic (QE), resonant, and deep inelastic (DIS) scattering. All current and forthcoming oscillation experiments live in the 0.1-10 GeV transition region, which encompasses the energy where the QE neutrino-nucleus interaction transitions into resonant scattering and the energy  where resonance scattering  transitions into DIS. 
 Neutrino and antineutrino QE charge current scattering refers to the process $\nu_\mu n \rightarrow \mu^- p$ and $\bar\nu_\mu p \rightarrow \mu^+ n$ where a charged lepton and single nucleon are ejected in the elastic interaction and the target nucleus remains at ground state. %of a neutrino (or antineutrino) with a nucleon in the target material.  
 Resonant scattering refers to an inelastic collision producing a nucleon excited state ($\Delta$,N*) -- the resonance, indeed -- which then quickly decays, most often to a nucleon and single-pion final state. DIS refers to the head-on collision between the neutrino and a parton inside the nucleon, producing hadronization and subsequent  abundant production of mesons and nucleons.  In addition to such interactions between the neutrino and a single component of the nucleus, neutrinos can also interact with the nucleus as a whole, albeit more rarely, a well documented process called coherent meson production scattering \cite{Vilain1993}; the signature of such process is the production of a distinctly forward-scattered single meson final state, most often a pion.  This simple picture of neutrino interactions works rather well for scattering 
off of light targets, such as the H$_2$ and D$_2$ of bubble chamber experiments \cite{RevModPhys.84.1307}, but the complexity of the nuclear structure for heavier nuclei such as argon complicates this model.

@@

Nuclear effects can potentially affect the neutrino event rates, nucleon emission, neutrino energy reconstruction, and neutrino/antineutrino ratios, carrying deep implications for oscillation experiments. 
Even in the case of ``simple" QE scattering, intra-nuclear hadron rescattering and correlation effects between the target nucleons can cause the ejection of additional nucleons in the final state, modifying the final state kinematics and topology. In case of resonant and DIS scattering, the hadronic interactions of meson and nucleons produced in the decay of the resonance or during hadronization  complicate this picture even more.
A large source of uncertainty in modeling  nuclear effects in neutrino interactions come from mesons interactions (and re-interactions) in the nucleus, e.g., pion re-scattering, charge exchange, and absorption.

A renewed interest for neutrino cross section measurements surged in recent year, along with lively discussion on the data reporting; the historical method of reporting the neutrino cross section as a function of the neutrino energy or momentum transferred shakes under the weight of its dependency on the chosen nuclear model.   On one hand, correcting for nuclear effects in neutrino interaction can introduce unwanted sources of uncertainty and model dependence especially due to the mis-modeling of the meson interactions. On the other, avoiding this correction makes a comparison between neutrino interactions on different target nuclei extremely difficult.

Data on neutrino scattering off many different nuclei are available for both charged current (CC) and neutral current (NC) channels, as summarized here \cite{RevModPhys.84.1307}. A summary of the results on QE, resonant and DIS scattering for neutrinos and antineutrinos from accelerators on different target is reported in Figure \ref{fig:nuXS}, where the (NUANCE) \cite{Casper2002} event generator is used as comparison with the theory. 
\begin{figure}[hbpt]
\centering
\includegraphics[scale=0.6]{Chapter-1/Images/NuCrossSec.png}
\includegraphics[scale=0.95]{Chapter-1/Images/AntiNuCrossSec.png}
\caption{Total neutrino (top) and antineutrino (bottom) CC cross sections per nucleon divided by neutrino energy as a function of energy as reported in \cite{RevModPhys.84.1307}. Predictions for the total (black), the QE (red), resonant (blue) and DIS (green) are provided by the NUANCE generator. The quasi-elastic scattering data and predictions have been averaged over neutron and proton targets (isoscalar target).}
\label{fig:nuXS}
\end{figure}
